SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-11-07 2017-Xudong-SMT/ActaInformatica.tex
proofreading and correction of new text...

Implementing these semantics raised several challenges:
\begin{itemize}
	\item to obtain a tool that could be applied to pNets representing
	various languages, in particular various action algebras,
	with their specific decision theories;
	\item to separate clearly the 
	algorithm 
	\YX{that generates the transitions of an open automaton from
	the symbolic reasoning part that uses an SMT
	engine to check the
	satisfiability of the predicates generated by the algorithm;}

@@

From a broader perspective, after the seminal work
of Hennessy, Lin, and their colleagues on value-passing systems with
assignments, many
different works addressing various classes of infinite-state systems
and/or parameterised topologies have been published, using
combinations of approaches, often including predicate abstraction and
SMT satisfiability (\eg \cite{DBLP:journals/jsat/AlbertiGPRR12,BruniEtAl-Tiles-Concur2000,CimattiEtAl-NUXMV-CAV2014,ChampionEtAl-Kind2-CAV2016,DBLP:conf/cade/GhilardiNRZ08}). 
Among these works, several have % shown the capability 
\YX{
demonstrated the usefulness
of SMT engines
(either Z3 or Yikes) as servers for solving verification conditions of
 algorithms, especially for large case-studies
(\eg
\cite{DBLP:journals/corr/abs-1806-11459,DBLP:journals/corr/CimattiGMT13}).
}
With
respect to these, we use symbolic representations not only to get a
finite representation of infinite spaces, but also to express the
(data-sensitive) synchronisations with the environment, making our
models suitable for compositional verification.

@@

\begin{example}[PLTS of the Controller in the ``Enable'' Operator]
\ERIC{
The PLTS of the controller $CState$ from Example \ref{example:EnableOperator} is  $\langle S,s_0, \to\rangle$, where:
\begin{itemize}
	\item[$\bullet$] $S=\{0, 1\}$, $s_0=0$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l}0, 0\xrightarrow{\delta}1, 1\xrightarrow{r}1\}$.
\end{itemize}
}
\QIN{
%Featuring guards and assignments in the label of transitions, 
\YX{The PLTS of the controller $CData$ is $\langle S',s'_0, \to\rangle$} where:
\begin{itemize}
	\item[$\bullet$] \YX{$S'=\{0\}$, $s'_0=0$, $vars(s'_0)=\{v\}$;} % $\texttt{init}\ \{v:=0\}$;
	\item[$\bullet$] $\to=\{0\xrightarrow{l, [v=0]}0, 0\xrightarrow{\delta,  [v=0], \{v\leftarrow1\}}0, 0\xrightarrow{r, [v=1]}0\}$.
\end{itemize}
}
\ERIC{The PLTSs contained in $EnableStateCompLeft$ are similar to the PLTS $CState$, but their action constants have been renamed, ensuring disjoint state variable sets.}
\end{example}

@@

\paragraph{Filtering.}
While matching a vector with a combination tuple, \emph{Matching}
tries to filter out some incompatibilities \QIN{in advance. If some of the subnet actions are marked inactive in the vector while some of the behaviours at the corresponding positions in the chosen combination are active, the matching would fail and the algorithm can filter it.}

@@

Our matching/filtering strategy builds some transitions where
the predicates express incompatible constraints. 
\QIN{Such constraints come from the mismatching of the actions by pattern-matching or unification, involving the guards, which are all presented in the predicates.}
Even if having an
unsatisfiable (symbolic) transition would not be incorrect,
we choose to minimize the open automaton (\ie its number of
transitions and states), by checking the predicates for satisfiability.

@@

\QIN{As the removal of the unsatisfiable transitions \YX{may entail some transitions without proceeding} transitions, }
the set $L1$ might contain unreachable transitions. To remove those transitions, the last step of the  
algorithm (lines 12-14) builds the final open automaton via a traversal of $L1$,
starting at the initial state $InitState(p)$ as defined in
page~\pageref{def-states}. More precisely, this is a residual
algorithm building a graph from the reachable transitions in $L1$. 
%and states are the set of target states of the transitions. 
Each state in the graph is decorated with a set of its local
variables, and a list of their possible assignments,
gathered during the traversal using the ``Post'' parts of the
OTs.

@@

In practice, this could be
expensive \QIN{as it needs several times of traversal of the whole set of open transitions}, and it is not incorrect to leave some UNSAT OTs in the
result. \QIN{\YX{Empirically, the unreachable transitions constitute  a small part of the ultimate} open transitions, very little will be gained to compute the fixed point. So we would only perform one step of iteration.} Further algorithms 
(e.g. model-checking or equivalence checking) will be able to use the
full information from the predicates and the assignment sets, even if
some spurious OTs are left here.

@@

The open automata generated for the two encodings of Enable in
Figure~\ref{schema:enable-pnets} are displayed in
Figure~\ref{schema:enable-OAs}. 
\QIN{The initial transition of $EnableData$ is transformed to the initialisation of the state variables, which is $\{v\leftarrow1\}$, for its open automata.}
Note that although these automata look very similar to the PLTS
controllers from Figure~\ref{schema:enable-pnets}, their nature is very
different: they represent the behavior of the whole system, including
the holes $P$ and $Q$. Their transitions do not rely on a particular
implementation of the controllers, but only encode the 
relations between the actions of the holes.
We have shown in \cite{henrio:Forte2016} 
\YX{that the two open automata are equivalent with respect to a notion of FH-bisimulation.}
%how to prove that they are bisimilar.

>>>>

Implementing these semantics raised several challenges:
\begin{itemize}
	\item to obtain a tool that could be applied to pNets representing
	various languages, in particular various action algebras,
	with their specific decision theories;
	\item to separate clearly the 
	algorithm that generates the transitions of an open automaton from
	the symbolic reasoning part that uses an SMT
	engine to check the
	satisfiability of the predicates generated by the algorithm;

@@

From a broader perspective, after the seminal work
of Hennessy, Lin, and their colleagues on value-passing systems with
assignments, many
different works addressing various classes of infinite-state systems
and/or parameterised topologies have been published, using
combinations of approaches, often including predicate abstraction and
SMT satisfiability (\eg \cite{DBLP:journals/jsat/AlbertiGPRR12,BruniEtAl-Tiles-Concur2000,CimattiEtAl-NUXMV-CAV2014,ChampionEtAl-Kind2-CAV2016,DBLP:conf/cade/GhilardiNRZ08}). 
Among these works, several have % shown the capability 
demonstrated the usefulness of SMT engines
(either Z3 or Yikes) as servers for solving verification conditions of
 algorithms, especially for large case-studies
(\eg
\cite{DBLP:journals/corr/abs-1806-11459,DBLP:journals/corr/CimattiGMT13}).
With
respect to these, we use symbolic representations not only to get a
finite representation of infinite spaces, but also to express the
(data-sensitive) synchronisations with the environment, making our
models suitable for compositional verification.

@@

\begin{example}[PLTS of the Controllers in the ``Enable'' Operator]

  \ERIC{
The PLTS of the controller $CState$ from Example \ref{example:EnableOperator} is  $\langle S,s_0, \to\rangle$, where:
$$S=\{0, 1\}; s_0=0;
	 \to=\{0\xrightarrow{l}0, 0\xrightarrow{\delta}1, 1\xrightarrow{r}1\}$$.
}
  \QIN{
%Featuring guards and assignments in the label of transitions, 
\YX{And the PLTS of the controller $CData$ is $\langle S',s'_0, \to\rangle$} where:
 \ERIC{$$S'=\{0\}; s'_0=0; vars(0)=\{v\}; \texttt{init}\ \{v:=0\};$$
	 $$\to=\{0\xrightarrow{l, [v=0]}0, 0\xrightarrow{\delta,  [v=0], \{v:=1\}}0, 0\xrightarrow{r, [v=1]}0\}$$.
  }}
\ERIC{The PLTSs contained in $EnableStateCompLeft$ are similar to the PLTS $CState$, but their action constants have been renamed, ensuring disjoint state variable sets.}
\end{example}

@@

\paragraph{Filtering.}
While matching a vector with a combination tuple, \emph{Matching}
tries to filter out some incompatibilities \ERIC{as early as possible.} \QIN{If some of the subnet actions are marked inactive in the vector while some of the behaviours at the corresponding positions in the chosen combination are active, the matching would fail and the algorithm can filter it.}

@@

Our matching/filtering strategy builds some transitions where
the predicates express incompatible constraints. 
\QIN{Such constraints come from the mismatching of the actions by pattern-matching or unification, \ERIC{including the guards collected from PLTSs and synchronisation vectors}, which are all present in the predicates.}
Even if having an
unsatisfiable (symbolic) transition would not be incorrect,
we choose to minimize the open automaton (\ie its number of
transitions and states), by checking the predicates for satisfiability.

@@

\QIN{As the removal of the unsatisfiable transitions \YX{may entail some transitions without preceeding} transitions, }
the set $L1$ might contain unreachable transitions. To remove those transitions, the last step of the  
algorithm (lines 12-14) builds the final open automaton via a traversal of $L1$,
starting at the initial state $InitState(p)$ as defined in
page~\pageref{def-states}. More precisely, this is a residual
algorithm building a graph from the reachable transitions in $L1$. 
%and states are the set of target states of the transitions. 
Each state in the graph is decorated with a set of its local
variables, and a list of their possible assignments,
gathered during the traversal using the ``Post'' parts of the
OTs.

@@

In practice, this could be
expensive \QIN{as it requires several traversals of the whole set of open transitions}, and it is not incorrect to leave some UNSAT OTs in the
result. \QIN{\YX{Empirically, on real life systems, the unreachable transitions constitute  a small part of the ultimate} open transitions, very little will be gained to compute the fixed point. So we would only perform one step of iteration.} Further algorithms 
(e.g. model-checking or equivalence checking) will be able to use the
full information from the predicates and the assignment sets, even if
some spurious OTs are left here.

@@

The open automata generated for the two encodings of Enable in
Figure~\ref{schema:enable-pnets} are displayed in
Figure~\ref{schema:enable-OAs}. 
\QIN{The initial transition of $EnableData$ is transformed to the initialisation of the state variables, which is $\{v\leftarrow 0\}$, for its open automata.}
Note that although these automata look very similar to the PLTS
controllers from Figure~\ref{schema:enable-pnets}, their nature is very
different: they represent the behavior of the whole system, including
the holes $P$ and $Q$. Their transitions do not rely on a particular
implementation of the controllers, but only encode the 
relations between the actions of the holes.
We have shown in \cite{henrio:Forte2016} 
\YX{that the two open automata are equivalent with respect to a notion of FH-bisimulation.}
%how to prove that they are bisimilar.
